W(ha/il)t's the (wo/ba)rd?

### 2022-04-30 paper_5.rmd
Add preambles

---
title: "W(ha/il)t's the (wo/ba)rd?"
thanks: >
  "What's the word?" or "Wilt's the bard?" (nonsense).
subtitle: "A cross-sectional analysis of English word pairs that satisfy the properties of the Split Decisions word puzzle"
author: "Oliver Daniel"
abstract: >
  From three English wordlist corpora, pairs of words are generated which might appear in the Split Decisions word puzzle as published by the New York Times, i.e., at least 5 characters in length and differing by exactly two consecutive characters. Even in absence of usage-related data, several ortho-phonotactic trends emerge that puzzle creators can employ to granularly increase or decrease the difficulty of guessing such a pair of words. Also discussed are morpheme and syllable boundaries, and how alterations of these boundaries within a word pair can make mentally recalling one given the other more difficult.
date: "April 30, 2022"
output:
  bookdown::pdf_document2:
    toc: true
    toc_depth: 1
    number_sections: false
knit: (function(input, ...) {
    rmarkdown::render(
    input,
    output_file = "out/paper_5.pdf",
    envir = globalenv(),
    ...
    )
  })
geometry: "margin=1.5in"
bibliography: in/references.bib
link-citations: true
---
\par \textbf{Keywords}: crosswords, split decisions, computational linguistics

@@

9. `type`: divides each word pair into one of three categories, depending on the positioning of its split pair. If the split pair is at the beginning of the word, the type is "prefix"; if at the end, "suffix"; and otherwise, "middle". (PL/GR)EASE is a "prefix"-type pair, for example.
10. `rank`: when comparing different pairs, splits, etc. by their frequency, a ranking variable is introduced to keep track of which is the most frequent, the second most frequent, and so on.
<!-- TODO -->

@@

\newpage
## Appendix
<!-- TODO: python script -->
### Link to Python preprocessing script
Found in the [GitHub repo (link)](https://github.com/oliver-daniel/inf_313_final_paper/blob/main/scripts/find_pairs.py).

>>>>

---
title: "W(ha/il)t's the (wo/ba)rd?"
thanks: >
  "What's the word?" or "Wilt's the bard?" (nonsense).
subtitle: "A cross-sectional analysis of English word pairs that satisfy the properties of the Split Decisions word puzzle"
author: "Oliver Daniel"
abstract: >
  From three English wordlist corpora, pairs of words are generated which might appear in the Split Decisions word puzzle as published by the New York Times, i.e., at least 5 characters in length and differing by exactly two consecutive characters. Even in absence of usage-related data, several ortho-phonotactic trends emerge that puzzle creators can employ to granularly increase or decrease the difficulty of guessing such a pair of words. Also discussed are morpheme and syllable boundaries, and how alterations of these boundaries within a word pair can make mentally recalling one given the other more difficult.
date: "April 30, 2022"
output:
  bookdown::pdf_document2:
    toc: true
    toc_depth: 1
    number_sections: false
knit: (function(input, ...) {
    rmarkdown::render(
    input,
    output_file = "out/paper_5.pdf",
    envir = globalenv(),
    ...
    )
  })
geometry: "margin=1.5in"
bibliography: in/references.bib
link-citations: true
---
<!-- PREAMBLE
# Purpose: Create all data analyses, tables, and visualizations required for the final paper of INF 312.
# Author: Oliver Daniel
# Data: 2022-04-30
# Contact: through Quercus
# License: WTFPL
# Pre-requisites: 
# - A reasonably modern R installation, and all data found in the `in/` directory of this repository. All other dependencies should auto-install on run.
# - knitr
 -->

\par \textbf{Keywords}: crosswords, split decisions, computational linguistics

@@

9. `type`: divides each word pair into one of three categories, depending on the positioning of its split pair. If the split pair is at the beginning of the word, the type is "prefix"; if at the end, "suffix"; and otherwise, "middle". (PL/GR)EASE is a "prefix"-type pair, for example.
10. `rank`: when comparing different pairs, splits, etc. by their frequency, a ranking variable is introduced to keep track of which is the most frequent, the second most frequent, and so on.

@@

\newpage
## Appendix
### Link to Python preprocessing script
Found in the [GitHub repo (link)](https://github.com/oliver-daniel/inf_313_final_paper/blob/main/scripts/find_pairs.py).

### 2022-04-30 scripts/exploration.rmd
Add preambles

```{r}
library(tidyverse)
library(stringdist)
```

@@

dat.all() |>
  select(word = x, split = split_x, start, end, corpus) |>
  rbind(dat.all() |>
    select(word = y, split = split_y, start, end, corpus)
  ) |>
  distinct() |>
  slice_sample(prop = .01) |>
  mutate(
    l_boundary = substr(word, start - 1, start - 1),
    split_l = substr(split, 1, 1),
    split_r = substr(split, 2, 2),
    r_boundary = substr(word, end + 1, end + 1)
  ) |>
  add_count(l_boundary, split_l, split_r, r_boundary) |>
  ggplot(aes(
    x = 0,
    xend = 1,
    y = match(l_boundary, letters) |> replace_na(27),
    yend = match(split_l, letters),
    color = corpus,
    size = n / (2 * max(n))
  )) +
  geom_segment(alpha = 0.2)
```

>>>>

<!-- PREAMBLE
# Purpose: Explorations and proto-visualizations of data for the final paper of INF 312.
# Author: Oliver Daniel
# Data: 2022-04-30
# Contact: through Quercus
# License: WTFPL
# Pre-requisites: 
# - A reasonably modern R installation, and all data found in the `in/` directory of this repository.
# - tidyverse
 -->

```{r}
library(tidyverse)
```

@@

dat.all() |>
  select(word = x, split = split_x, start, end, corpus) |>
  rbind(dat.all() |>
    select(word = y, split = split_y, start, end, corpus)
  ) |>
  mutate(
    l_boundary = substr(word, start - 1, start - 1),
    split_l = substr(split, 1, 1),
    split_r = substr(split, 2, 2),
    r_boundary = substr(word, end + 1, end + 1)
  ) |>
  count(l_boundary, split_l, split_r, r_boundary, corpus) |>
  filter(n > 250) |>
  ggplot(aes(
    x = 0,
    xend = 1,
    y = match(l_boundary, rev(letters)) |> replace_na(27),
    yend = match(split_l, rev(letters)),
    color = corpus,
    size = n / (max(n)^2),
    alpha = (n / max(n) - .1)^3 + 0.2
  )) +
  geom_segment() +
  scale_y_discrete(breaks = c(rev(letters), "[prefix]"),
    sec.axis = sec_axis(~ .)
  )
```
